Probabilities of polar aligned disks in binary star systems

### 2024-08-21 src/tex/ms.tex
Move analytic part to earlier

\DeclareMathOperator{\arctantwo}{arctan2}

@@

@@

\citet{martin2019} give two conditions for libration, each valid for a different regime of constant of motion $\chi$ (see their Equations 31, 32, 38).
\TJ{(Would it be helpful to define $\chi$?)}
These conditions can be rearranged to find the minimum value of $\Omega$ that will allow for a librating orbit.
\begin{equation}
    \label{eq:omega_min}
    \begin{aligned}
        \sin^2{\Omega_{\rm min}} & = \frac{(1-e_{\rm b}^2) (2j + \cos^2{i})}{5 e_{\rm b}^2 \sin^2{i}} & \text{if} ~\chi \ge 0 \\
    & = \frac{2}{5 \sin^2{i}} + \frac{\cos{i}}{5j \sin^2{i}} - \frac{e_{\rm b}^2}{20 j^2 (1-e_{\rm b}^2) \sin^2{i}} & \text{if} ~\chi < 0
    \end{aligned}
\end{equation}

And in the $\chi < 0$ case $i_{\rm crit}$ satisfies
\begin{equation}
    \label{eq:icrit2}
    \sin^2{i_{\rm crit}} - \frac{\cos{i_{\rm crit}}}{5j} + \frac{e_{\rm b}^2}{20 j^2 (1-e_{\rm b}^2)} + \frac{2}{5} = 0
\end{equation}

Such that if the LHS of Equation \ref{eq:icrit1} or \ref{eq:icrit2} (depending on the value of $\chi$) evaluate to $<0$ then only circulating orbits are allowed.

To get the probability of a polar orbit given randomly distributed $\Omega$ we can use a modified version of Equation 8 from \citet{ceppi2024}.
\begin{equation}
    \begin{aligned}
        P(j,e_{\rm b}, i) & = 1 - \frac{2}{\pi} \Omega_{\rm min} & \text{if} ~i > i_{\rm crit} \\
        & = 0 & \text{if} ~i \le i_{\rm crit}
    \end{aligned}
\end{equation}

From which the fraction of total polar orbits can be calculated
\begin{equation}
    \label{eq:fp_int}
    f_{\rm polar} = \iiint P(j,e_{\rm b}, i)\, dj\,de_{\rm b}\,di
\end{equation}

This integral is straightforward to compute numerically, and allows for any desired distribution in $j$, $e_{\rm b}$, or $i$, so long as they are normalized.
For example, an isotropic distribution of disks would require we replace $di$ with $\frac{1}{2}\sin{u}\,du$ for $u \in [0,\pi]$.

\subsubsection{High-$j$ limit}
\label{subsubsec:hi_j}
When $j$ is very large, the $\chi < 0$ branch of Equation \ref{eq:omega_min} becomes
\begin{equation}
    \sin^2{\Omega_{\rm min}} = \frac{2}{5}\csc^2{i}
\end{equation}

which has no dependence on the disk angular momentum or on the binary eccentricity. We can then find that the polar fraction in this limit is
\begin{equation}
    f_{{\rm polar}, j \gg 1} = \int_0^\pi\,\left (1-\frac{2}{\pi}\arcsin{\left (\sqrt{\frac{2}{5}} \csc{i} \right )}\right ) \sin{i}\,di = \variable{output/high_j_integral.txt}
\end{equation}

>>>>

\DeclareMathOperator{\arctantwo}{arctan2}
\DeclareMathOperator{\arccsc}{arccsc}

@@

\section{Analytic Treatment}
\label{sec:analytic}
For a massive ring misaligend to its binary host stars, \citet{martin2019} give two conditions for libration. Each condition valid for a different regime of constant of motion $\chi$ (see their Equations 31, 32, 38), defined as
\begin{equation}
    \chi = e_{\rm b}^2 - 2\,(1 - e_{\rm b}^2)\,j\,(2j + \cos{i})
\end{equation}
for binary eccentricity $e_{\rm b}$, normalized ring angular momentum $j = J_{\rm r}/J_{\rm b}$, and mutual inclination $i$. These conditions can be rearranged to find the minimum value of $\Omega$ that will allow for a librating orbit.
\begin{equation}
    \label{eq:omega_min}
    \sin^2{\Omega_{\rm min}} = 
    \left \{
    \begin{array}{lr}
         \frac{(1-e_{\rm b}^2) (2j + \cos^2{i})}{5 e_{\rm b}^2 \sin^2{i}} & \text{if} ~\chi \ge 0 \\
    \frac{2}{5 \sin^2{i}} + \frac{\cos{i}}{5j \sin^2{i}} - \frac{e_{\rm b}^2}{20 j^2 (1-e_{\rm b}^2) \sin^2{i}} & \text{if} ~\chi < 0
    \end{array}
    \right .
\end{equation}

And in the $\chi < 0$ case $i_{\rm crit}$ satisfies
\begin{equation}
    \label{eq:icrit2}
    \sin^2{i_{\rm crit}} - \frac{\cos{i_{\rm crit}}}{5j} + \frac{e_{\rm b}^2}{20 j^2 (1-e_{\rm b}^2)} + \frac{2}{5} = 0
\end{equation}
Such that if the LHS of Equation \ref{eq:icrit1} or \ref{eq:icrit2} (depending on the value of $\chi$) evaluate to $<0$ then only circulating orbits are allowed.

To get the probability of a polar orbit given randomly distributed $\Omega$ we can use a modified version of Equation 8 from \citet{ceppi2024}.
\begin{equation}
    P(j,e_{\rm b}, i) = 
    \left \{
    \begin{array}{lr}
         1 - \frac{2}{\pi} \Omega_{\rm min} & \text{if} ~i > i_{\rm crit} \\
        0 & \text{if} ~i \le i_{\rm crit}
    \end{array}
    \right .
\end{equation}
From which the fraction of total polar orbits can be calculated
\begin{equation}
    \label{eq:fp_int}
    f_{\rm polar} = \iiint P(j,e_{\rm b}, i)\, dj\,de_{\rm b}\,di
\end{equation}
This integral is straightforward to compute numerically, and allows for any desired distribution in $j$, $e_{\rm b}$, or $i$, so long as they are normalized. For example, an isotropic distribution of disks would require we replace $di$ with $\frac{1}{2}\sin{u}\,du$ for $u \in [0,\pi]$.

\subsection{High-$j$ limit}
\label{subsec:high_j}

When $j$ is very large, the $\chi < 0$ branch of Equation \ref{eq:omega_min} becomes
\begin{equation}
    \sin^2{\Omega_{\rm min}} = \frac{2}{5}\csc^2{i}
\end{equation}

The value of $i_{\rm crit}$ is then
\begin{equation}
    i_{\rm crit} = \arccsc{\sqrt{5/2}} = 39.2^\circ
\end{equation}
which is the inclination of the last circular orbit for Kozai-Lidov oscillations in the limit that the perturbed object is very far from the perturbers \citep{kozai1962,lidov1962}. It also has no dependence on the disk angular momentum or on the binary eccentricity. For an isotropic angular momentum distribution, find that the polar fraction in this limit is
\begin{equation}
    f_{{\rm polar}, j \gg 1} = \int_0^\pi\,\left (1-\frac{2}{\pi}\arcsin{\left (\sqrt{\frac{2}{5}} \csc{i} \right )}\right ) \sin{i}\,di = \variable{output/high_j_integral.txt}
\end{equation}

@@
